---
title: "Time Line and Map of Earthquakes for package: earthquakes"
author: "Christian Raggi"
date: "5/23/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Line and Map of Earthquakes for package: earthquakes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Load package 'earthquakes'

```{r load package}
library(earthquakes)
```

## How to read the data into the R environment?

Run the following code:

```{r earthquakes dataset}
filename <- system.file("extdata", "signif.txt", package = "earthquakes")
library(readr)
eq_data_raw <- readr::read_delim(file = filename, delim = "\t")
eq_data <- eq_clean_data(eq_raw = eq_data_raw)
head(eq_data_raw$LOCATION_NAME)
head(eq_data$LOCATION_NAME)
```

## How to make a time line plot of earthquakes for different countries?

See the R code below, for example:

```{r time line, fig.width=7}
library(dplyr)
library(ggplot2)
library(ggthemes)
eq_data %>%
    dplyr::filter(DATE >= "1995-01-01" & 
                      DATE <= "2018-01-01" &
                      COUNTRY %in% c("USA","ITALY","MEXICO")) %>%
    ggplot(aes(x = DATE,
               y = COUNTRY,
               colour = DEATHS,
               size = EQ_PRIMARY)) +
    geom_timeline() +
    geom_timeline_label(aes(x = DATE, 
                            y = COUNTRY, 
                            tags = LOCATION_NAME, 
                            number = 3, 
                            max_aes = EQ_PRIMARY)) +
    theme_tufte() +
    theme(legend.position = "right")
```


## How to map the earthquakes?

Each earthquake should be shown with a circle, and the radius of the circle should be proportional to the earthquake's magnitude (EQ_PRIMARY). Your code, assuming you have the earthquake data saved in your working directory as "earthquakes.tsv.gz", should be able to be used in the following way:

```{r map, fig.width=7}
library(dplyr)
library(lubridate)
library(leaflet)
eq_data %>%
    filter(COUNTRY %in% "ITALY" &
               lubridate::year(DATE) >= 2000) -> eq_data_sample

eq_data_sample %>%
    eq_clean_data() %>%
    filter(COUNTRY %in% "ITALY" & lubridate::year(DATE) >= 2000) %>%
    eq_map(annot_col = "DATE")

eq_data_sample %>%
    eq_clean_data() %>%
    filter(COUNTRY %in% "ITALY" & lubridate::year(DATE) >= 2000) %>%
    mutate(popup_text = eq_create_label(.)) %>%
    eq_map(annot_col = "popup_text")
```

